### 2021-07-08 rad-icfp-2021/intro.tex
some fixes

% TODO: rewrite for more theory people. some ideas: for canonical than Plotkin's thing because normal semantics, complexity proof compared to all other recent papers, 
% all papers that have implementations, their interface in the paper is too high level and misses a lot of things (e.g. in \cite{vytiniotis2019differentiable,lantern_icfp,sherman2021}) compared to implementation, hinders progress. 
% For instance, in \cite{lantern_icfp} it's a lot of manually optimized and hand-derived primitives to achieve good performance.

@@

    \begin{center}
            \begin{tabular}{l}
                !let w$_1$ = x$_1$ * x$_2$ in!\\ 
                !let w$_2$ = w$_1$ * x$_1$ in!\\
                !let y$_1$,y$_2$,y$_3$,y$_4$,y$_5$ = 0,0,0,0,1 in!\\
                !let y$_1'$! != y$_1$+w$_1$*y$_5$ in!\\
                !let y$_4'$! != y$_4$+x$_1$*y$_5$ in!\\
                !(y$_1'$+x$_2$*y$_4'$!!, y$_2$+x$_1$*y$_4'$!!, y$_3$)!
            \end{tabular}
    \end{center}   
This is equivalent to the SSA form \cite{cytron1989efficient} of what the imperative reverse-mode differentiation of our initial term would be.
This term can be further optimized via constant propagation and algebraic simplifications to give
        \begin{center}
            \begin{tabular}{{c}}
                !let w$_1$ = x$_1$ * x$_2$ in!\\ 
                !let w$_2$ = w$_1$ * x$_1$ in!\\
                !(w$_1$+x$_2$*x$_1$, x$_1$*x$_1$, 0)!
            \end{tabular}
        \end{center}
    \end{example}

@@

% Our contribution makes these claims precise. In more details, we introduce
% \begin{itemize}
%     \item an expressive source language with array operations, yet restrictive enough for efficient differentiation (\S\ref{sub:sourcelang})
%     \item a method for extending our work and modular approaches to add and prove correct new primitives as needed in practice (\S\ref{sec:generalization})
%     \item a new intermediate representation (IR), UNF, essentially turning every operation into a unary one which inputs and outputs a tuple (\S\ref{sec:unf})
%     \item a transformation between the IR and the main language (\S\ref{sub:transformations to and from UNF})
%     \item a simple reverse-mode transformation on the IR (\S\ref{sub:Simple reverse mode transformation})
%     \item the direct reverse-mode transformation obtained from the source to the target language (\S\ref{sub:Macro for pure reverse mode transformation})
%     \item a denotational semantics of our languages and transformations using multicategories and concategories (\S\ref{sec:correctness})
%     \item correctness of the transformations (\S\ref{sec:correctness})
%     \item complexity guarantees (\S\ref{sec:complexity})
%     \item how efficiency is ensured via standard optimization techniques (\S\ref{sub:Optimizations})
%     \item several additions to the source language (\S\ref{sec:generalization}) TODO:should emphasize that it should be easier for us to add primitives compared to several recent work. give recipe.
% \end{itemize}

>>>>

@@

    \begin{center}
            \begin{tabular}{l}
                !let w$_1$ = x$_1$ * x$_2$ in!\\ 
                !let w$_2$ = w$_1$ * x$_1$ in!\\
                !let y$_1$,y$_2$,y$_3$,y$_4$,y$_5$ = 0,0,0,0,1 in!\\
                !let y$_1'$! != y$_1$+w$_1$*y$_5$ in!\\
                !let y$_4'$! != y$_4$+x$_1$*y$_5$ in!\\
                !(y$_1'$+x$_2$*y$_4'$!!, y$_2$+x$_1$*y$_4'$!!, y$_3$)!
            \end{tabular}
    \end{center}   
This is very close to the SSA form \cite{cytron1989efficient} of what the imperative reverse-mode differentiation of our initial term would be.
This term can be further optimized via constant propagation and algebraic simplifications to give
        \begin{center}
            \begin{tabular}{{c}}
                !let w$_1$ = x$_1$ * x$_2$ in!\\ 
                !let w$_2$ = w$_1$ * x$_1$ in!\\
                !(w$_1$+x$_2$*x$_1$, x$_1$*x$_1$, 0)!
            \end{tabular}
        \end{center}
    \end{example}

@@

% Our contribution makes these claims precise. In more details, we introduce
% \begin{itemize}
%     \item an expressive source language with array operations, yet restrictive enough for efficient differentiation (\S\ref{sub:sourcelang})
%     \item a method for extending our work and modular approaches to add and prove correct new primitives as needed in practice (\S\ref{sec:generalization})
%     \item a new intermediate representation (IR), UNF, essentially turning every operation into a unary one which inputs and outputs a tuple (\S\ref{sec:unf})
%     \item a transformation between the IR and the main language (\S\ref{sub:transformations to and from UNF})
%     \item a simple reverse-mode transformation on the IR (\S\ref{sub:Simple reverse mode transformation})
%     \item the direct reverse-mode transformation obtained from the source to the target language (\S\ref{sub:Macro for pure reverse mode transformation})
%     \item a denotational semantics of our languages and transformations using multicategories and concategories (\S\ref{sec:correctness})
%     \item correctness of the transformations (\S\ref{sec:correctness})
%     \item complexity guarantees (\S\ref{sec:complexity})
%     \item how efficiency is ensured via standard optimization techniques (\S\ref{sub:Optimizations})
%     \item several additions to the source language (\S\ref{sec:generalization})
% \end{itemize}
